Limits to Ecological Forecasting: Estimating Uncertainty for Critical Transitions with Deep Learning

### 2022-08-16 manuscript/manuscript.Rmd
Reverting one edit

Examining patterns in the scores over time (Fig 6-7) provides a more nuanced understanding of the forecast dynamics than aggregate scores alone (Fig 2).
In the Hopf bifurcation, CRPS scores get worse over time across all methods, including the MCMC forecasts.
In the saddle node bifurcation and stochastic transition, the same pattern holds somewhat more dramatically for non-MCMC forecast, while MCMC scores are worse in mid-range.
Comparing CRPS scores to logs score also emphasizes the relative role of uncertainty: for instance, the MCMC scores for the Hopf bifurcation get steadily worse under CRPS but not under logs score. 
A relatively sharp transition can be seen under both MCMC scores on the Hopf bifurcation once the magnitude of the oscillations exceeds the variance created by mere stochasticity: the MCMC model no longer over-estimates the spread of the data, while the ML models now underestimate that variation. 
CRPS scores for stochastic transitions exhibit a distinct two-branch pattern, with scores for a given replicate being either very high (poor skill) or very low, reflecting whether the individual 'true' replicate matches the mean state predicted by the forecast or the other state. 
Logs skill score may be a better measure in this context, where correctly capturing the uncertainty in the forecast means that this bi-modal structure in scores can be avoided entirely, e.g. by the MCMC predictions. 
The forecast-skill-over time plots illustrate different reasons for the bi-modal distribution in skill seen for the saddle-node and stochastic transition scenarios in Fig 2 respectively: in the case of the saddle node, the two modes are distinguished by time-horizon: short term forecasts are relatively accurate, and longer term forecasts (i.e. after the catastrophic transition) are poor.
In the case of the stochastic transition model, the two modes are not structured by horizon but by replicate, with some replicates having transitioned and others still in the original state.

>>>>

Examining patterns in the scores over time (Fig 6-7) provides a more nuanced understanding of the forecast dynamics than aggregate scores alone (Fig 2).
In the Hopf bifurcation, CRPS scores get worse over time across all methods, including the MCMC forecasts.
In the saddle node bifurcation and stochastic transition, the same pattern holds somewhat more dramatically for non-MCMC forecast, while MCMC scores are worst in mid-range.
Comparing CRPS scores to logs score also emphasizes the relative role of uncertainty: for instance, the MCMC scores for the Hopf bifurcation get steadily worse under CRPS but not under logs score. 
A relatively sharp transition can be seen under both MCMC scores on the Hopf bifurcation once the magnitude of the oscillations exceeds the variance created by mere stochasticity: the MCMC model no longer over-estimates the spread of the data, while the ML models now underestimate that variation. 
CRPS scores for stochastic transitions exhibit a distinct two-branch pattern, with scores for a given replicate being either very high (poor skill) or very low, reflecting whether the individual 'true' replicate matches the mean state predicted by the forecast or the other state. 
Logs skill score may be a better measure in this context, where correctly capturing the uncertainty in the forecast means that this bi-modal structure in scores can be avoided entirely, e.g. by the MCMC predictions. 
The forecast-skill-over time plots illustrate different reasons for the bi-modal distribution in skill seen for the saddle-node and stochastic transition scenarios in Fig 2 respectively: in the case of the saddle node, the two modes are distinguished by time-horizon: short term forecasts are relatively accurate, and longer term forecasts (i.e. after the catastrophic transition) are poor.
In the case of the stochastic transition model, the two modes are not structured by horizon but by replicate, with some replicates having transitioned and others still in the original state.
